---
title: "Report 6"
author: "Nathan Bana"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "svg", message=FALSE, warning=FALSE)
```
This report uses the following R packages:
```{r packages}
library(tidyverse)
library(knitr)
library(dslabs)
```
We will be using the `gapminder` dataset from the `dslabs` package. This dataset contains health and income outcomes for 184 countries from 1960 to 2016. More precisely, it contains the following 9 variables:

* country.
* year.
* infant_mortality. Infant deaths per 1000.
* life_expectancy. Life expectancy in years.
* fertility. Average number of children per woman.
* population. Country population.
* gpd. GDP according to World Bankdev.
* continent.
* region. Geographical region.

Let's take a look at the data types:
```{r str}
str(gapminder)
```
We can see that there are 5 numeric vectors, 3 factors and 1 integer vector.
The question I want to answer is whether the average life expectancy is different between African and European countries for the year 2016. First, let's filter the `gapminder` dataset so that it only includes observations of African and European countries for the year 2016:
```{r filter}
gapminder2 <- gapminder %>%
  filter(continent %in% c("Africa","Europe"), year==2016)
str(gapminder2)
```
Now, we have what we need to create a plot of the average life expectancy by continent:
```{r dot_plot, fig.asp=148/665.95, fig.fullwidth=TRUE}
ggplot(gapminder2, aes(life_expectancy, continent)) + 
  stat_summary(fun = mean, geom = "point") +
  scale_x_continuous(limits = c(0,90),
                     breaks = seq(0,90,10)) +
  labs(y = "Continent",
       x = "Average life expectancy (years)",
       title = "Average life expectancy by continent for the year 2016.",
       subtitle = "Only Africa and Europe are included.",
       caption = "Source: Gapminder") +
  coord_fixed(3) +
  theme_classic() +
  theme(
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    rect = element_rect(fill = "#adebeb"),
    panel.background = element_rect(fill = "#c1f0f0")
  )
```

We can see that the average life expectancy is clearly different between Europe and Africa. Furthermore, given the large effect size and sample size, we can already be pretty sure that this difference is statistically significant. But let's still run a hypothesis test, just for fun (woo-hoo!).

To compare 2 means, we would have to run a t-test. First, let's visually check the normality assumption:
```{r hist}
ggplot(gapminder2, aes(life_expectancy)) +
  geom_histogram(binwidth = 5,
                 center=2.5) +
  facet_grid(cols = vars(continent))
```

Uh-oh! Africa looks fine, but Europe is another story. Let's run a Shapiro-Wilk test to be sure:
```{r shapiro}
gapminder_europe <- gapminder %>%
  filter(continent == "Europe", year==2016)
shapiro.test(gapminder_europe$life_expectancy)
```
Yikes. The test is significant, meaning the data are very probably not normally distributed. Fortunately, we can just use a nonparametric test like the Wilcoxon test to compare our 2 means:
```{r wilco}
wilcox.test(life_expectancy ~ continent, gapminder2)
```
The test is significant (*p*=8.21e-15), meaning we can safely reject the null hypothesis that the population means are identical. We can conclude that the difference in means between the two continents is not due to chance.
