---
title: "Report 8"
author: "Nathan Bana"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "svg")
```
This report uses the following R packages:
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(devtools)
library(ggbiplot)
```
In this report we will use the famous `iris` dataset. First, let's run a PCA:
```{r pca}
iris.pca <- prcomp(iris[, 1:4], scale = F)
summary(iris.pca)
ggscreeplot(iris.pca)
```

We see that the first 2 PCs explain almost the entirety of the variance (over 97%). Therefore, we should only keep those 2. Let's run a biplot:
```{r biplot}
ggbiplot(iris.pca, choices = 1:2, alpha = 0.1)
```

We can clearly see 2 distinct groups of observations. We can also note that petal length and petal width are very strongly positively correlated.

Next, let's add the PC scores to the original dataset:
```{r scores}
iris.scores <- cbind(iris, iris.pca$x[, 1:2])
```
Now we can plot the PCs:
```{r plot}
ggplot(iris.scores, aes(PC1, PC2, color = Species)) +
  geom_point() +
  coord_fixed() +
  theme_classic()
```

Thanks to the coloring, we now see that there actually are 3 distinct groups of observations.
