---
title: "Report 4"
author: "Nathan Bana"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This report uses the following R packages:
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(dslabs)
```
We will be using the `gapminder` dataset from the `dslabs` package. This dataset contains health and income outcomes for 184 countries from 1960 to 2016. More precisely, it contains the following 9 variables:

* country.
* year.
* infant_mortality. Infant deaths per 1000.
* life_expectancy. Life expectancy in years.
* fertility. Average number of children per woman.
* population. Country population.
* gpd. GDP according to World Bankdev.
* continent.
* region. Geographical region.

Let's take a look at the data types:
```{r str}
str(gapminder)
```
We can see that there are 5 numeric vectors, 3 factors and 1 integer vector.
Let's transform the dataset into long format:
```{r long}
gapminder_long <- gapminder %>% 
  pivot_longer(c("infant_mortality", "life_expectancy", "fertility", "population", "gdp"), names_to = "variable")
gapminder_long %>% head() %>% kable()
```
Now, let's unite the continent and region columns:
```{r unite}
gapminder_long <- gapminder_long %>% 
  unite(cont_region, continent, region)
gapminder_long %>% head() %>% kable()
```
Actually, I regret everything. Let's revert the change:
```{r separate}
gapminder_long <- gapminder_long %>% 
  separate(cont_region, into=c("continent", "region"), sep="_")
gapminder_long %>% head() %>% kable()
```
Finally, using our original `gapminder` dataset, let's plot infant mortality against life expectancy:
```{r plot, warning=FALSE}
ggplot(gapminder, aes(infant_mortality, life_expectancy)) +
  geom_point()
```

