---
title: "Report 5"
author: "Nathan Bana"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```
This report uses the following R packages:
```{r packages}
library(tidyverse)
library(knitr)
library(dslabs)
library(scales)
```
We will be using the `gapminder` dataset from the `dslabs` package. This dataset contains health and income outcomes for 184 countries from 1960 to 2016. More precisely, it contains the following 9 variables:

* country.
* year.
* infant_mortality. Infant deaths per 1000.
* life_expectancy. Life expectancy in years.
* fertility. Average number of children per woman.
* population. Country population.
* gpd. GDP according to World Bankdev.
* continent.
* region. Geographical region.

Let's take a look at the data types:
```{r str}
str(gapminder)
```
We can see that there are 5 numeric vectors, 3 factors and 1 integer vector.

## Fertility by region

I want to know the distribution of fertility by region for the year 2015, arranged in ascending order by the median. Our first step is filtering the data to include only the year 2015.
```{r filter}
gapminder2015 <- filter(gapminder, year == 2015)
kable(head(gapminder2015))
str(gapminder2015)
```
Now we can use this filtered data set to create our plot:
```{r plot1}
ggplot(gapminder2015, aes(reorder(region, fertility, na.rm = T, FUN = median), fertility, fill=continent)) +
  geom_boxplot(outlier.alpha = 0.3, outlier.size = 1) +
  scale_y_continuous(limits = c(1,8),
                     breaks = seq(1,8,1)) +
  labs(x = "Region",
       y = "Fertility (average number of children per woman)",
       fill = "Continent",
       title = "Fertility by region for the year 2015.",
       subtitle = "Each observation represents a country.\nOne country was excluded due to missing data.",
       caption = "Source: Gapminder") +
  coord_flip() +
  theme_classic() +
  theme(
    axis.title.y = element_text(angle = 0, vjust = 0.5)
  )
```

It is clear from the plot that African countries tend to have higher fertility than Western countries.

## Life expectancy vs. population

For our next plot, I want to explore the relationship between life expectancy and population for each continent. I will focus on 6 specific years: 1960, 1970, 1980, 1990, 2000 and 2010. Our first step is filtering the data to include only the years of interest.
```{r filter2}
gapminder1960_2010 <- filter(gapminder, year %in% c("1960", "1970", "1980", "1990", "2000", "2010"))
kable(head(gapminder1960_2010))
str(gapminder1960_2010)
```
Now we can use this filtered data set to create our plot:
```{r plot2}
ggplot(gapminder1960_2010, aes(population/10^6, life_expectancy, color = continent)) +
  geom_point(size = 1, alpha = 0.5) +
  scale_x_continuous(trans='log10',
                     breaks=trans_breaks('log10', function(x) 10^x),
                     labels=trans_format('log10', math_format(10^.x))) +
  labs(x = "Population in millions (log scale)",
       y = "Life expectancy (years)",
       color = "Continent",
       title = "Life expectancy vs. population (1960-2010).",
       subtitle = "Each data point represents a country.",
       caption = "Source: Gapminder") +
  facet_wrap(vars(year), ncol = 2) +
  theme_classic() +
  theme(
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    panel.grid.major = element_line(color = "grey"),
    strip.background = element_rect(fill = "#ffcccc")
  )
```

We can see that there does not seem to be a relationship between population and life expectancy. It is nevertheless interesting to see that from 1960 to 2010, life expectancy increased considerably almost everywhere on the planet. We can also see that Europe has always had the highest life expectancy, and Africa the lowest.
