---
title: "Report 10"
author: "Nathan Bana"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "svg")
```
This report uses the following R packages:
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(dslabs)
```
# Dataset

We will be using the `gapminder` dataset from the `dslabs` package. This dataset contains health and income outcomes for 184 countries from 1960 to 2016. More precisely, it contains the following 9 variables:

* country.
* year.
* infant_mortality. Infant deaths per 1000.
* life_expectancy. Life expectancy in years.
* fertility. Average number of children per woman.
* population. Country population.
* gpd. GDP according to World Bankdev.
* continent.
* region. Geographical region.

Let's take a look at the data types:
```{r str}
str(gapminder)
```
We can see that there are 5 numeric vectors, 3 factors and 1 integer vector. For this report about time series analysis, I am only interested in the evolution of life expectancy across time in Switzerland. Let us filter the data accordingly:
```{r filter}
gapminder.filtered <- gapminder %>% 
  filter(country == "Switzerland") %>% 
  select(year, life_expectancy)
str(gapminder.filtered)
kable(head(gapminder.filtered))
```
# Analysis

Now, let us create a time series object with our data and plot it:
```{r ts}
gapminder.ts <- ts(gapminder.filtered$life_expectancy, 1960, 2016)
ts.plot(gapminder.ts, xlab = "Year", ylab = "Life expectancy (years)", 
        main = "Life expectancy in Switzerland, 1960-2016")
```

We can observe that life expectancy has been steadily increasing in Switzerland since 1960.

Now let us plot the autocorrelation:
```{r acf}
acf(gapminder.ts)
```

We can see that the autocorrelation is very positively strong overall, meaning that the time series is strongly dependent on its past, and that when life expectancy is high on any given year, it tends to be also high the preceding year (for lag 1).

Next, let us fit the autoregressive (AR) model to our data and try to predict the future of our time series:
```{r AR}
gapminder.ar <- arima(gapminder.ts, order = c(1, 0, 0))
predict(gapminder.ar, n.ahead = 10)
```
Hmm. The model predicts that life expectancy will go down in Switzerland from 2017 to 2026. It is very curious given the data that we fed it...

Let us try fitting the simple moving average (MA) model to our data instead, and try to predict the future with that:
```{r MA}
gapminder.ma <- arima(gapminder.ts, order = c(0, 0, 1))
predict(gapminder.ma, n.ahead = 10)
```
Well now the results are even weirder...
